---
title: "final_project_working_copy"
author: "Matthew Lawlor"
date: "11/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidyquant)
library(scales)
```

## Read and tidy

### H1N1

***Kaggle***

```{r}
h1n1_df = read.csv("./data/h1n1_data.csv") %>% 
  janitor::clean_names() %>% 
  select(-link) %>% 
  mutate(
    date = as.Date(date, "%m/%d/%y"),
    pandemic_interval = "2009/04/24" %--% date,
    pandemic_duration = as.duration(pandemic_interval) / ddays(1)
  ) %>% 
 rename(h1n1_cumulative_cases = cumulative_no_of_cases, h1n1_cumulative_deaths = cumulative_no_of_deaths) %>% 
  arrange(pandemic_duration)
  
h1n1_global = 
  h1n1_df %>%   
  group_by(pandemic_duration) %>% 
  summarize(
    h1n1_cumulative_cases = sum(h1n1_cumulative_cases, na.rm = TRUE),
    h1n1_cumulative_deaths = sum(h1n1_cumulative_deaths, na.rm = TRUE)
  ) %>% 
  mutate(
    h1n1_new_cases = h1n1_cumulative_cases - lag(h1n1_cumulative_cases),
    h1n1_new_deaths = h1n1_cumulative_deaths - lag(h1n1_cumulative_deaths)
  )

save(h1n1_global, file = "h1n1_global.rda")

h1n1_us = 
  h1n1_df %>% 
  filter(country == "United States of America") %>% 
  mutate(
    h1n1_new_cases = h1n1_cumulative_cases - lag(h1n1_cumulative_cases),
    h1n1_new_deaths = h1n1_cumulative_deaths - lag(h1n1_cumulative_deaths)
  )
```

***WHO***

```{r}
##US data
h1n1_us_who = readxl::read_excel("./data/USA_FluNetInteractiveReport.xlsx", skip = 5) %>% 
  janitor::clean_names() %>% 
  select(start_date:end_date, h1n1_new_cases = a_h1n1_pdm09) %>% 
  mutate(
    start_date = as.Date(start_date),
    end_date = as.Date(end_date),
    pandemic_interval = "2009/04/06" %--% end_date,
    pandemic_duration = as.duration(pandemic_interval) / ddays(1)
  ) %>% 
    filter(pandemic_duration >= 0) %>% 
  mutate(
    h1n1_new_cases = as.numeric(h1n1_new_cases),
    h1n1_cumulative_cases = CUMULATIVE_SUM(h1n1_new_cases)
  )

##global data
h1n1_global_who = read.csv("./data/global_FluNetInteractiveReport.csv", skip = 3) %>% 
  janitor::clean_names() %>% 
  select(country, sdate, edate, h1n1_new_cases = ah1n12009) %>% 
  mutate(
    sdate = as.Date(sdate, "%m/%d/%y"),
    edate = as.Date(edate, "%m/%d/%y"),
    pandemic_interval = "2009/04/06" %--% edate,
    pandemic_duration = as.duration(pandemic_interval) / ddays(1)
  ) %>% 
    filter(pandemic_duration >= 0) %>% 
  arrange(pandemic_duration) %>% 
  group_by(pandemic_duration) %>% 
  summarize(
    h1n1_new_cases = sum(h1n1_new_cases, na.rm = TRUE)
  ) %>% 
  mutate(
    h1n1_cumulative_cases = CUMULATIVE_SUM(h1n1_new_cases)
  )
```


### COVID-19

***US***

```{r}
covid_us_cases = read_csv("./data/case_daily_trends_united_states.csv", skip = 3) %>% 
  janitor::clean_names() %>% 
  select(-x7_day_moving_avg) %>% 
  mutate(
    date = as.Date(date, "%b %d %Y"),
    pandemic_interval = "2020/01/22" %--% date,
    pandemic_duration = as.duration(pandemic_interval) / ddays(1)
  ) %>% 
  rename(covid_new_cases = new_cases) %>% 
  arrange(pandemic_duration) %>% 
  mutate(
    covid_cumulative_cases = CUMULATIVE_SUM(covid_new_cases)
  )

covid_us_deaths = read.csv("./data/death_daily_trends_united_states.csv", skip = 3) %>%
  janitor::clean_names() %>% 
  select(-x7_day_moving_avg) %>%
  mutate(
    date = as.Date(date, "%b %d %Y"),
    pandemic_interval = "2020/01/22" %--% date,
    pandemic_duration = as.duration(pandemic_interval) / ddays(1)
  ) %>% 
  rename(covid_new_deaths = new_deaths) %>% 
  arrange(pandemic_duration) %>% 
  mutate(
    covid_cumulative_deaths = CUMULATIVE_SUM(covid_new_deaths)
  )
```

***Global***

```{r}
covid_global = read_csv("./data/covid19_global.csv") %>% 
  janitor::clean_names() %>% 
  select(-country_code, -who_region, covid_new_cases = new_cases, covid_cumulative_cases = cumulative_cases, covid_new_deaths = new_deaths, covid_cumulative_deaths = cumulative_deaths) %>% 
  mutate(
    date = as.Date(date_reported, "%m/%d/%Y"),
    pandemic_interval = "2020/01/03" %--% date,
    pandemic_duration = as.duration(pandemic_interval) / ddays(1)
  ) %>% 
  group_by(pandemic_duration) %>% 
  summarize(across(covid_new_cases:covid_cumulative_deaths, sum))

save(covid_global, file = "covid_global.rda")
```


### SARS

```{r}
sars_us = read.csv("./data/sars_2003_complete_dataset_clean.csv") %>% 
  janitor::clean_names() %>% 
  select(-number_recovered, sars_cumulative_cases = cumulative_number_of_case_s, sars_cumulative_deaths = number_of_deaths) %>%
  filter(country == "United States") %>%
  mutate(
    date = as.Date(date),
    pandemic_interval = "2003/03/17" %--% date,
    pandemic_duration = as.duration(pandemic_interval) / ddays(1),
    sars_new_cases = sars_cumulative_cases - lag(sars_cumulative_cases),
    sars_new_deaths = sars_cumulative_deaths - lag(sars_cumulative_deaths)
  )

save(sars_us, file = "sars_us.rda")

sars_global = read.csv("./data/sars_2003_complete_dataset_clean.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    date = as.Date(date),
    pandemic_interval = "2003/03/17" %--% date,
    pandemic_duration = as.duration(pandemic_interval) / ddays(1)
  ) %>% 
    select(pandemic_duration, sars_cumulative_cases = cumulative_number_of_case_s, sars_cumulative_deaths = number_of_deaths) %>% 
  group_by(pandemic_duration) %>% 
  summarize(
    sars_cumulative_cases = sum(sars_cumulative_cases),
    sars_cumulative_deaths = sum(sars_cumulative_deaths)
  ) %>% 
  mutate(
    sars_new_cases = sars_cumulative_cases - lag(sars_cumulative_cases),
    sars_new_deaths = sars_cumulative_deaths - lag(sars_cumulative_deaths)
  )

<<<<<<< HEAD

save(sars_us, file = "sars_us.rda")

=======
save(sars_us, file = "sars_us.rda")
>>>>>>> da024c34662f8951f125b965f7e7321f9e2c400e
save(sars_global, file = "sars_global.rda")
```

## Visualization

### US H1N1

#### Kaggle cases

***New cases***
```{r}
h1n1_us %>%
  ggplot(aes(x = pandemic_duration, y = h1n1_new_cases)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1)
```

***Cumulative cases***
```{r}
 h1n1_us %>%
  ggplot(aes(x = pandemic_duration, y = h1n1_cumulative_cases)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1)
```

#### WHO cases

***New cases***
```{r}
h1n1_us_who %>%
  ggplot(aes(x = pandemic_duration, y = h1n1_new_cases)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
    labs(
    title = "H1N1: new cases, United States",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

***Cumulative cases***
```{r}
 h1n1_us_who %>%
  ggplot(aes(x = pandemic_duration, y = h1n1_cumulative_cases)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
  labs(
    title = "H1N1: Cumulative cases, United States",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

#### Deaths

***New deaths***
```{r}
 h1n1_us %>% ##kaggle data
  ggplot(aes(x = pandemic_duration, y = h1n1_new_deaths)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
    labs(
    title = "H1N1: New deaths, United States",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

***Cumulative deaths***
```{r}
h1n1_us %>% ##kaggle data
  ggplot(aes(x = pandemic_duration, y = h1n1_cumulative_deaths)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
    labs(
    title = "H1N1: Cumulative deaths, United States",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

### US COVID-19

***New cases***

```{r}
covid_us_cases %>% 
  ggplot(aes(x = pandemic_duration, y = covid_new_cases)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
    labs(
    title = "COVID-19: new cases, United States",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

***Cumulative cases***
```{r}
covid_us_cases %>% 
  ggplot(aes(x = pandemic_duration, y = covid_cumulative_cases)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
    labs(
    title = "COVID-19: cumulative cases, United States",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

***New deaths***

```{r}
covid_us_deaths %>% 
  ggplot(aes(x = pandemic_duration, y = covid_new_deaths)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
    labs(
    title = "COVID-19: new deaths, United States",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

***Cumulative deaths***

```{r}
covid_us_deaths %>% 
  ggplot(aes(x = pandemic_duration, y = covid_cumulative_deaths)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
    labs(
    title = "COVID-19: cumulative deaths, United States",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

### US SARS

***New cases***

```{r}
sars_us %>% 
  ggplot(aes(x = pandemic_duration, y = sars_new_cases)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
    labs(
    title = "SARS: new cases, United States",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

***Cumulative cases***
```{r}
 sars_us %>%
  ggplot(aes(x = pandemic_duration, y = sars_cumulative_cases)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
    labs(
    title = "SARS: cumulative cases, United States",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

***New deaths***

```{r}
sars_us %>% 
  ggplot(aes(x = pandemic_duration, y = sars_new_deaths)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
    labs(
    title = "SARS: new deaths, United States",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

***Cumulative deaths***

```{r}
sars_us %>% 
  ggplot(aes(x = pandemic_duration, y = sars_cumulative_deaths)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
    labs(
    title = "SARS: cumulative deaths, United States",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

### Merge US cases

```{r} 
##using WHO cases
us_cases =
  left_join(covid_us_cases, h1n1_us_who, by = "pandemic_duration")

us_cases =
  full_join(us_cases, sars_us, by = "pandemic_duration") %>% 
  select(pandemic_duration, h1n1_new_cases, h1n1_cumulative_cases, covid_new_cases, covid_cumulative_cases, sars_new_cases, sars_cumulative_cases) %>% 
    arrange(pandemic_duration) %>% 
  pivot_longer(
    h1n1_new_cases:sars_cumulative_cases,
    names_to = "class",
    values_to = "n"
  ) %>% 
  separate(class, into = c("disease", "trend"), sep = "_", extra = "merge")
```

***US New cases***

```{r}
##linear
us_cases %>% 
  filter(trend == "new_cases") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n =7, size = 1) +
  labs(
    title = "New cases, United States",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Linear plot"
  )
```

```{r}
##log-transformed
us_cases %>% 
  filter(trend == "new_cases") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n =7, size = 1) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x), labels = trans_format("log10", math_format(10^.x))) +
  labs(
    title = "New cases, United States",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Log-transformed plot"
  )
```

***US Cumulative cases***

```{r}
##linear
us_cases %>% 
  filter(trend == "cumulative_cases") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n =7, size = 1) +
  labs(
    title = "Cumulative cases, United States",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Linear plot"
  )
```

```{r}
##log-transformed
us_cases %>% 
  filter(trend == "cumulative_cases") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n =7, size = 1) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x), labels = trans_format("log10", math_format(10^.x))) +
  labs(
    title = "Cumulative cases, United States",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Log-transformed plot"
  )
```

### Merge US deaths

```{r}
us_deaths = ##using Kaggle deaths
  full_join(h1n1_us, covid_us_deaths, by = "pandemic_duration")

us_deaths =
  full_join(us_deaths, sars_us) %>% 
  select(pandemic_duration, h1n1_new_deaths, h1n1_cumulative_deaths, covid_new_deaths, covid_cumulative_deaths, sars_new_deaths, sars_cumulative_deaths) %>% 
  arrange(pandemic_duration) %>% 
  pivot_longer(
    h1n1_new_deaths:sars_cumulative_deaths,
    names_to = "class",
    values_to = "n"
  ) %>% 
  separate(class, into = c("disease", "trend"), sep = "_", extra = "merge")
```

***New deaths***

```{r}
##linear
us_deaths %>% 
  filter(trend == "new_deaths") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
  labs(
    title = "New deaths, United States",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Linear plot"
  )
```

```{r}
##log-transformed
us_deaths %>% 
  filter(trend == "new_deaths") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x), labels = trans_format("log10", math_format(10^.x))) +
  labs(
    title = "New deaths, United States",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Log-transformed plot"
  )
```

***Cumulative deaths***

```{r}
##linear
us_deaths %>% 
  filter(trend == "cumulative_deaths") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
  labs(
    title = "Cumulative deaths, United States",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Linear plot"
  )
```

```{r}
#log-transformed
us_deaths %>% 
  filter(trend == "cumulative_deaths") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x), labels = trans_format("log10", math_format(10^.x))) +
  labs(
    title = "Cumulative deaths, United States",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Log-transformed plot"
  )
```

### Global H1N1

***New cases***
```{r}
h1n1_global_who %>%
  ggplot(aes(x = pandemic_duration, y = h1n1_new_cases)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
  labs(
    title = "H1N1: new cases, global",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

***Cumulative cases***
```{r}
 h1n1_global_who %>%
  ggplot(aes(x = pandemic_duration, y = h1n1_cumulative_cases)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
  labs(
    title = "H1N1: cumulative cases, global",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

### Global COVID

***New cases***

```{r}
covid_global %>% 
  ggplot(aes(x = pandemic_duration, y = covid_new_cases)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
      labs(
    title = "COVID-19: new cases, global",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

***Cumulative cases***
```{r}
covid_global %>% 
  ggplot(aes(x = pandemic_duration, y = covid_cumulative_cases)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
      labs(
    title = "COVID-19: cumulative cases, global",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

***New deaths***

```{r}
covid_global %>% 
  ggplot(aes(x = pandemic_duration, y = covid_new_deaths)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
        labs(
    title = "COVID-19: new deaths, global",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

***Cumulative deaths***

```{r}
covid_global %>% 
  ggplot(aes(x = pandemic_duration, y = covid_cumulative_deaths)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
        labs(
    title = "COVID-19: cumulative cases, global",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

### Global SARS

***New cases***

```{r}
sars_global %>% 
  ggplot(aes(x = pandemic_duration, y = sars_new_cases)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
        labs(
    title = "SARS: new cases, global",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

***Cumulative cases***
```{r}
sars_global %>% 
  ggplot(aes(x = pandemic_duration, y = sars_cumulative_cases)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
    labs(
    title = "SARS: cumulative cases, global",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

***New deaths***

```{r}
sars_global %>% 
  ggplot(aes(x = pandemic_duration, y = sars_new_deaths)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
        labs(
    title = "SARS: new deaths, global",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

***Cumulative deaths***

```{r}
sars_global %>% 
  ggplot(aes(x = pandemic_duration, y = sars_cumulative_deaths)) +
  geom_point(alpha = .5) +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
        labs(
    title = "SARS: cumulative deaths, global",
    x = "Pandemic Duration, days",
    y = "n"
  )
```

### Merge global cases

```{r} 
##using WHO cases
global_cases =
  full_join(covid_global, h1n1_global_who, by = "pandemic_duration")

global_cases =
  full_join(global_cases, sars_global, by = "pandemic_duration") %>% 
  select(pandemic_duration, h1n1_new_cases, h1n1_cumulative_cases, covid_new_cases, covid_cumulative_cases, sars_new_cases, sars_cumulative_cases) %>% 
  arrange(pandemic_duration) %>% 
  pivot_longer(
    h1n1_new_cases:sars_cumulative_cases,
    names_to = "class",
    values_to = "n"
  ) %>% 
  separate(class, into = c("disease", "trend"), sep = "_", extra = "merge")
```

***Global new cases***

```{r}
##linear
global_cases %>% 
  filter(trend == "new_cases") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
  labs(
    title = "New cases, global",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Linear plot"
  )
```

```{r}
##log-transformed
global_cases %>% 
  filter(trend == "new_cases") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x), labels = trans_format("log10", math_format(10^.x))) +
  labs(
    title = "New cases, global",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Log-transformed plot"
  )
```

***Global cumulative cases***

```{r}
##linear
global_cases %>% 
  filter(trend == "cumulative_cases") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
  labs(
    title = "Cumulative cases, global",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Linear plot"
  )
```

```{r}
##log-transformed
global_cases %>% 
  filter(trend == "cumulative_cases") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x), labels = trans_format("log10", math_format(10^.x))) +
  labs(
    title = "Cumulative cases, global",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Log-transformed plot"
  )
```

### Merge global deaths

```{r}
global_deaths = ##using Kaggle deaths
  full_join(h1n1_us, covid_global, by = "pandemic_duration")

global_deaths =
  full_join(global_deaths, sars_global) %>% 
  select(pandemic_duration, h1n1_new_deaths, h1n1_cumulative_deaths, covid_new_deaths, covid_cumulative_deaths, sars_new_deaths, sars_cumulative_deaths) %>% 
  arrange(pandemic_duration) %>% 
  pivot_longer(
    h1n1_new_deaths:sars_cumulative_deaths,
    names_to = "class",
    values_to = "n"
  ) %>% 
  separate(class, into = c("disease", "trend"), sep = "_", extra = "merge")
```

***New deaths***

```{r}
##linear
global_deaths %>% 
  filter(trend == "new_deaths") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
  labs(
    title = "New deaths, global",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Linear plot"
  )
```

```{r}
##log-transformed
global_deaths %>% 
  filter(trend == "new_deaths") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x), labels = trans_format("log10", math_format(10^.x))) +
  labs(
    title = "New deaths, global",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Log-transformed plot"
  )
```

***Cumulative deaths***

```{r}
##linear
global_deaths %>% 
  filter(trend == "cumulative_deaths") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
  labs(
    title = "Cumulative deaths, global",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Linear plot"
  )
```

```{r}
#log-transformed
global_deaths %>% 
  filter(trend == "cumulative_deaths") %>% 
  ggplot(aes(x = pandemic_duration, y = n, color = disease)) +
  geom_point() +
  geom_ma(ma_fun = SMA, n = 7, size = 1) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x), labels = trans_format("log10", math_format(10^.x))) +
    labs(
    title = "Cumulative deaths, global",
    x = "Pandemic Duration, days",
    y = "n",
    caption = "Log-transformed plot"
  )
```
